---
title: "GridBot Model"
output:
  html_notebook: default
  pdf_document: default
---
#GridBot
##Lindsey Sample and Josh DeLeew, 2019

Grid world:
A matrix of rows and colums, each square has a light value, can be occupied by an agent, and can have an obstacle
2 layers of matrix: light and obstacle + agent
light: 0 = dark, 255= max light (or 0-1)

light scaline function: 255*exp(-distance*.5)
falls off, if we multiply d by a number less than one, drops off slower
more than 1, drops off faster

Here, we make our light grid and obstacle grid.
If 1 is entered, it will make a static environment. If 2, dynamic
Dynamic picks a row and column number that = 100 total squares and places 5 random obstacles

```{r}
makeLightGrid<- function(numRows, numCols){
  lightGrid<- matrix(NA, numRows, numCols)
  for(x in 1:numCols){
    for(y in 1:numRows){
      #puts in middle of grid
      distance <- sqrt((0-y)^2+((numCols+1)/2-x)^2)
      lightGrid[y,x] <- round(255*exp(-distance*.5))
    }
  }
  return (lightGrid)
}

#takes in 1 or 2 (static or dynamic)
makeGrids<- function(x){
  #if static, 10x10
  if (x == 1){
    numRows <- 10
    numCols <- 10
  }
  #if dynamic, pick from dynamic nums list
  if (x==2){
    d<- dynamicNums()
  numRows <- d[1]
  numCols <- d[2]
  }
  #make light grid with appropriate dimesions, add barriers along edges
  lightGrid<- makeLightGrid(numRows, numCols)
  obsGrid<- matrix(NA, numRows, numCols)
  obsGrid[1,]<- 1
  obsGrid[,1]<- 1
  obsGrid[numRows,]<- 1
  obsGrid[,numCols] <-1
  #if static, plus barrier in center
  if(x==1){
    obsGrid[round(numRows/2), round(numCols/2)]<- 1
    obsGrid[round(numRows/2)+1, round(numCols/2)]<- 1
    obsGrid[round(numRows/2), round(numCols/2)+1]<- 1
    obsGrid[round(numRows/2)-1, round(numCols/2)]<- 1
    obsGrid[round(numRows/2), round(numCols/2)-1]<- 1
  } 
  #if dynamic, placing 5 obstacles randomly
  if (x==2){
    for(i in 1:6){
    r<- sample(numRows-1, 1)
    c<- sample(numCols-1, 1)
    obsGrid[r, c]<- 1
    }
  }
  #place bot in spot right above barrier in center (rounded to avoid decimal problems)
    obsGrid[numRows-1, round(numCols/2)]<- 0
    return(list(obsGrid, lightGrid))
}
#gets random numbers that make a 100 square grid
dynamicNums<-function(){
  #cant have 100x1 and 2x50, no obstacle options
  listy<- list( c(20,5), c(5, 20), c(10,10), c(4, 25), c(25, 4))
  listNums<- (sample(listy, 1))
  pairNums<- listNums[[1]]
  nRows<- pairNums[[1]]
  nCols <- pairNums[[2]]
  return (c(nRows, nCols))
}



```

Agent traverses block matrix
blank indicates no block
1 indicates block
0 = agent is in this spot
agent cannot be on a block


functions:
(determined by NN)
turn in place to left, turn to right, step forward, step backward

move:
takes state as argument
changes grid or cardinal so agent is in new spot
return new grid

moveForward: moves in the direction bot is facing one space forward

cardinal: an ordered pair that designates next move
y, x
(rows, col)
(1,0) -> down
(0, 1) -> right
(-1,0) -> up
(0, -1) -> left
```{r}
moveForward <-function(state){
  cardinal <- state[[1]]
  obsGrid <- state[[2]]
  
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  if(is.na(obsGrid[aCur+cardinal])){
    obsGrid[aCur]<- NA
    obsGrid[aCur+cardinal]<- 0
  }
  state[[2]] <- obsGrid
  return(state)
}
moveBackward <-function(state){
  cardinal <- state[[1]]
  obsGrid <- state[[2]]
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  if(is.na(obsGrid[aCur-cardinal])){
    obsGrid[aCur]<- NA
    obsGrid[aCur-cardinal]<- 0
  }
  state[[2]] <- obsGrid
   return(state)
}

turnClock <- function(state){
  cardinal <- state[[1]]
  obsGrid <- state[[2]]
  #if down, turn left
  if(all(cardinal == c(1,0))){
    cardinal <- c(0,-1)
  }
  #if left, turn up
  else if(all(cardinal == c(0,-1))){
    cardinal <- c(-1,0)
  }
  #if up, turn right
  else if(all(cardinal == c(-1,0))){
    cardinal <- c(0,1)
  }
   #if right, turn down
  else if(all(cardinal == c(0,1))){
    cardinal <- c(1,0)
  }
  state[[1]] <- cardinal
  return(state)
}
turnCounter <- function(state){
  cardinal <- state[[1]]
  obsGrid <- state[[2]]
  #if left, turn down
  if(all(cardinal == c(0,-1))){
    cardinal <- c(1,0)
  }
  #if up turn left
  else if(all(cardinal == c(-1,0))){
    cardinal <- c(0,-1)
  }
  #if right, turn up
  else if(all(cardinal == c(0,1))){
    cardinal <- c(-1,0)
  }
   #if down, turn right
  else if(all(cardinal == c(1,0))){
    cardinal <- c(0,1)
  }
  state[[1]] <- cardinal
  return(state)
}

```

#light function
get light takes in obsGrid, finds the position of the bot (aCur)
based on the cardinal, it takes the average of the space its in, the space in front, and the two to the right and left for the right and left sensors respectively. The sum of r and l is the number added to the accumulator

! [LDR] (LDR.jpg)


```{r}
down <- c(1,0)
right<- c(0, 1)
up<- c(-1,0)
left<- c(0, -1)

rightLight<- 0
leftLight<- 0

getRightLight<- function(state){
  cardinal <- state[[1]]
  obsGrid <- state[[2]]
  lightGrid <-state[[3]]
  
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  #up
  if(all(cardinal== up)){
    rightLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+right], lightGrid[aCur+up+right])
       }
  
  #right
  else if(all(cardinal== right)){
      rightLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+up], lightGrid[aCur+up+right])
       }
  #left
    else if(all(cardinal== left)){
    rightLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+up], lightGrid[aCur+up+left])
      }
  #down
    else if(all(cardinal== down)){
    rightLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+right], lightGrid[aCur+down+right])
     }
  
 
  
  return(rightLight)
}

getLeftLight<- function(state){
  cardinal <- state[[1]]
  obsGrid <- state[[2]]
  lightGrid <-state[[3]]
  
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  #up
  if(all(cardinal== up)){
  leftLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+left], lightGrid[aCur+up+left])
  }
  
  #right
  else if(all(cardinal== right)){
     leftLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+down], lightGrid[aCur+down+right])
  }
  #left
    else if(all(cardinal== left)){
leftLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+down], lightGrid[aCur+down+left])
    }
  #down
    else if(all(cardinal== down)){
   leftLight<- mean(lightGrid[aCur], lightGrid[aCur+cardinal], lightGrid[aCur+left], lightGrid[aCur+down+left])
  }
  
  return(leftLight)
}


```

###IRs
returns the distance of the closest block

```{r}
leftIR<- 0
rightIR<-0

getRightIR<- function(state){
    cardinal <- state[[1]]
  obsGrid <- state[[2]]
  lightGrid <-state[[3]]
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  #up
  #checks for both Right  IR, they include the two spaces in front, sets to distance between,
  # first checks before, then to right, then up up, then up up right
  #else if because we check in order
  
  
  #up direction
  if(all(cardinal== up)){ 
    #if item in spot 1 (up), IR = distance
    if(!is.na(obsGrid[aCur+up])){
      rightIR <- dist(rbind(aCur, aCur + up))[1]
      #if item in spot 2 (right), IR = distance
    } else if (!is.na(obsGrid[aCur+up+right])){
      rightIR <- dist(rbind(aCur, aCur + up + right))[1]
      #if item in spot 3, IR= distance
    } else if (!is.na(obsGrid[aCur+up+up])){
      rightIR <- dist(rbind(aCur, aCur + up + up))[1]
      #if item in spot 4, IR = distance
    } else if (!is.na(obsGrid[aCur+up+up+right])){
      rightIR <- dist(rbind(aCur, aCur + up + up + right))[1]
    }}
  #down direction
  if (all(cardinal == down)){
    #spot 1
    if(!is.na(obsGrid[aCur+down])){
         rightIR <- dist(rbind(aCur, aCur + down))[1]
         #spot 2
    } else if (!is.na(obsGrid[aCur+down+right])){
       rightIR <- dist(rbind(aCur, aCur + down + right))[1]
       #spot 3
    } else if (!is.na(obsGrid[aCur+down+down])){
    rightIR <- dist(rbind(aCur, aCur + down + down))[1]
    #spot 4
    } else if (!is.na(obsGrid[aCur+down+down+right])){
      rightIR <- dist(rbind(aCur, aCur + down + down + right))[1]
    }
  }
  
  #right direction
  if (all(cardinal == right)){
    #spot 1
    if(!is.na(obsGrid[aCur+right])){
        rightIR <- dist(rbind(aCur, aCur + right))[1]
         #spot 2
    } else if (!is.na(obsGrid[aCur+down+right])){
       rightIR <- dist(rbind(aCur, aCur + down + right))[1]
       #spot 3
    } else if (!is.na(obsGrid[aCur+right+right])){
    rightIR <- dist(rbind(aCur, aCur + right + right))[1]
    #spot 4
    } else if (!is.na(obsGrid[aCur+down+right+right])){
      rightIR <- dist(rbind(aCur, aCur + down + right + right))[1]
    }
  }
  
  if (all(cardinal == left)){
    #spot 1
    if(!is.na(obsGrid[aCur+left])){
         rightIR <- dist(rbind(aCur, aCur + left))[1]
         #spot 2
    } else if (!is.na(obsGrid[aCur+up+left])){
       rightIR <- dist(rbind(aCur, aCur + up + left))[1]
       #spot 3
    } else if (!is.na(obsGrid[aCur+left+left])){
    rightIR <- dist(rbind(aCur, aCur + left + left))[1]
    #spot 4
    } else if (!is.na(obsGrid[aCur+up+left+left])){
      rightIR <- dist(rbind(aCur, aCur + up + left + left))[1]
    }
  }
  
  return (abs(rightIR))
}
    
    
getLeftIR<- function(state){
    cardinal <- state[[1]]
  obsGrid <- state[[2]]
  lightGrid <-state[[3]]
  
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  if(all(cardinal== up)){ 
    #if item in spot 1 (up), IR = distance
    if(!is.na(obsGrid[aCur+up])){
      leftIR <- dist(rbind(aCur, aCur + up))[1]
      #if item in spot 2 (left), IR = distance
    } else if (!is.na(obsGrid[aCur+up+left])){
      leftIR <- dist(rbind(aCur, aCur + up + left))[1]
      #if item in spot 3, IR= distance
    } else if (!is.na(obsGrid[aCur+up+up])){
     leftIR <- dist(rbind(aCur, aCur + up + up))[1]
      #if item in spot 4, IR = distance
    } else if (!is.na(obsGrid[aCur+up+up+left])){
      leftIR <- dist(rbind(aCur, aCur + up + up + left))[1]
    }}
   if (all(cardinal == down)){
    #spot 1
    if(!is.na(obsGrid[aCur+down])){
         leftIR <- dist(rbind(aCur, aCur + down))[1]
         #spot 2
    } else if (!is.na(obsGrid[aCur+down+left])){
       leftIR <- dist(rbind(aCur, aCur + down + left))[1]
       #spot 3
    } else if (!is.na(obsGrid[aCur+down+down])){
    leftIR <- dist(rbind(aCur, aCur + down + down))[1]
    #spot 4
    } else if (!is.na(obsGrid[aCur+down+down+left])){
      leftIR <- dist(rbind(aCur, aCur + down + down + left))[1]
    }
   }
  #right
  if (all(cardinal == right)){
    #spot 1
    if(!is.na(obsGrid[aCur+right])){
         leftIR <- dist(rbind(aCur, aCur + right))[1]
         #spot 2
    } else if (!is.na(obsGrid[aCur+up+right])){
       leftIR <- dist(rbind(aCur, aCur + right))[1]
       #spot 3
    } else if (!is.na(obsGrid[aCur+right+right])){
    leftIR <- dist(rbind(aCur, aCur + right + right))[1]
    #spot 4
    } else if (!is.na(obsGrid[aCur+up+right+right])){
      leftIR <- dist(rbind(aCur, aCur + up + right + right))[1]
    }
  }

     if (all(cardinal == left)){
    #spot 1
    if(!is.na(obsGrid[aCur+left])){
        leftIR <- dist(rbind(aCur, aCur + left))[1]
         #spot 2
    } else if (!is.na(obsGrid[aCur+down+left])){
       leftIR <- dist(rbind(aCur, aCur + down + left))[1]
       #spot 3
    } else if (!is.na(obsGrid[aCur+left+left])){
   leftIR <- dist(rbind(aCur, aCur + left + left))[1]
    #spot 4
    } else if (!is.na(obsGrid[aCur+down+left+left])){
      leftIR <- dist(rbind(aCur, aCur + down + left + left))[1]
    }
  }   
 
  return(abs(leftIR))
}
  
```
###Bumper
returns 1 if there is an object in front of the bot (depends on where it is facing)
```{r}
bump <- 0
getBump<- function(state){
      cardinal <- state[[1]]
  obsGrid <- state[[2]]
  lightGrid <-state[[3]]
  
  aCur<- which(obsGrid==0, arr.ind=TRUE)
  if(is.na(obsGrid[aCur+cardinal])){
    bump<-0
  }
  else if(!is.na(obsGrid[aCur+cardinal])){
    bump<-1
  }
  

  return(bump)
}
```


###Update sensors
```{r}
updateInput<- function(state){
    leftIR<- getLeftIR(state)
    rightIR<- getRightIR(state)
    leftLight<- getLeftLight(state)
    rightLight<- getRightLight(state)
    bump<- getBump(state)
    
    input<- c(leftLight,leftIR, bump, rightIR, rightLight) 
    return(input)
}

```

###MakeMove
```{r}
makeMove<- function(weights, state){
  cardinal <- state[[1]]
  obsGrid <- state[[2]]
  lightGrid <-state[[3]]
  
  input<- updateInput(state)
  
  output<- c(0,0,0,0)
  #weights<- matrix(NA,length(input), length(output))
  
  #matrix and input vector, matrix multiplication (dot product of every column)
  #OR for loop, dot product (sum, vector * vector)
  
  #print(weights)
  
  for (i in 1:length(output)){
      output[i]<- 1 / (1 + exp(-sum(input*weights[,i])))
    
  }
  sum.output<- sum(output)
  adj.output <- output/sum.output
  
  
  action <- sample(c(moveForward,moveBackward, turnClock, turnCounter),1,  prob=adj.output)
  new.state <- action[[1]](state)
  
  state <- new.state
  
  
  
#Update State: Takes data of gen's run in Grid, Gives next Genome
  if(plotState){
    p <- gridBotANN.PlotState(state)
    ggsave(paste0('plots/',plot.num,'.png'), p, device = png())
    plot.num <<- plot.num + 1
    dev.off()
  }
  
  return(new.state)
}


```
###Making weights and setting up
```{r}
makeRandWeights <- function(){
  output<- c(moveForward, moveBackward, turnClock, turnCounter)
  weight.options<- c(-1, -.5, 0, .5, 1)
  weights<- matrix(NA,5, length(output))
  for (row in 1:nrow(weights)){
    for (col in 1:ncol(weights)){
    weights[row,col] <- sample(weight.options, 1)
    }
  }
  return (weights)
}


setUp<- function(x){
  grids<- makeGrids(x)
  obsGrid <- grids[[1]]
  lightGrid <- grids[[2]]
  state <- list(
    cardinal = c(-1,0),
    obsGrid,
    lightGrid
  )
  #weights <- makeRandWeights()
  return(state)
}

# t<-setUp(2)
# 
# test1.state <- setUp(2)
# weights <- makeRandWeights()
# for(i in 1:100){
#   test1.state <- makeMove(weights,test1.state)
# }


```

upate sensor vals

```{r}
# N.moves <-100
# N.bots <- 10
# N.gen <- 100

#runs bots with 100 moves

runBot <- function(weights, x){

  # make a starting environment / state
  state <- setUp(x)
  total.light <- 0
  for(i in 1:N.moves){
    state <- makeMove(weights, state)
    total.light <- total.light + getLeftLight(state) + getRightLight(state)
  }
  
  return(total.light)
}

#runBot(makeRandWeights(), 2)
  
runGeneration <- function(genomes, x) {
  fitness.vals <- data.frame(bot.number=1:10, fitness=rep(0,10))
  for(i in 1:N.bots){
   #print(genomes[[i]])
    #print(i)
    fitness <- runBot(genomes[[i]], x)
    fitness.vals[i, 'fitness'] <- fitness
   
    }
  return(fitness.vals)
  # runBot for each genome in the data frame
}


createGeneration <- function(N.bots){
  genomes <- list()
  for(i in 1:N.bots){
    #setUp(10,10)
    genomes[[i]] <- makeRandWeights()    
  }
  return(genomes)
}

# test1.state <- setUp(10, 10)
# weights <- makeRandWeights()
# for(i in 1:100){
#   test1.state <- makeMove(weights,test1.state)
# }





#takes in what runGeneration returns
next.genomes<- function(old.gen, old.genomes){
  
  next.generation <- sample(old.gen$bot.number, 10, replace=T, prob = old.gen$fitness)
  #next.generation <- c(1,1,1,1,4,5,6,3,3,4)
  
  next.bots.genomes<- old.genomes
  
  for (i in 1:length(next.generation)) {
    #print(i)
    bot<- next.generation[[i]]
    next.bots.genomes[i] <- old.genomes[bot]
  }
  #next.bots.genomes <- old.genomes[next.generation]
  return(next.bots.genomes)
}

#r<-nrow(weights)
  
# r<- as.numeric(sample(r, 1))


  w<-makeRandWeights()
  weight.options<- c(-1, -.5, 0, .5, 1)
 w[1, 1] <- as.numeric(sample(weight.options, 1))
 w

mutate <- function(weights){
 
   
  new.weights <- weights
  
  weight.options<- c(-1, -.5, 0, .5, 1)
  #i<- sample(length(weights), 1)
  #dim(weights)
  
  r<-as.numeric(nrow(weights))
  c<-as.numeric(ncol(weights))
  
  r.m<- as.numeric(sample(r, 1))
  c.m<- as.numeric(sample(c, 1))

  #print(new.weights)
  #print(new.weights[r.m, c.m])
  new.weights[r.m, c.m] <- as.numeric(sample(weight.options, 1))
  
  #print(new.weights)
  return(new.weights)
}

  
  #new.weights <- rnorm(length(weights), mean=weights, sd=0.05)
  #10% mutation, change weights

#skip that do a loop change one

#for loop, 1 to total gens, does all steps

#mutates genomes
# for (i in 1:length(next.bots)){
#    next.bots[[i]] <- mutate(next.bots[[i]])
# }


```

```{r}

# run<- function(x){
#   #x<-1
#   genomes<- createGeneration(N.bots)
#   gen<- runGeneration(genomes, x)
# 
#   for (a in 2:N.gen) {
#    print(a)
#     if(a == N.gen){
#       #plotState <<- T
#     } else {
#       plotState <<- F
#     }
#     genomes<- next.genomes(gen, genomes)
#      for (i in 1:length(genomes)) {
#        #print(genomes[i])
#        #i<-1
#       genomes[[i]]<- mutate(genomes[[i]])
#       #print(genomes[i])
#     }
#     gen <- runGeneration(genomes, x)
# 
#   }
# 
#   top.scorer <- as.numeric(which.max(gen$fitness))
#   
#   
#   final.weights<-genomes[[top.scorer]]
#   return(final.weights)
# }

run<- function(x){
  #x<-1
  genomes<- createGeneration(N.bots)
  gen<- runGeneration(genomes, x)

  gen.history <- gen
  gen.history$generation <- 1
  
  for (a in 2:N.gen) {
   print(a)
    if(a == N.gen-1){
       plotState <- T
    } else {
      plotState <- F
    }
    genomes<- next.genomes(gen, genomes)
     for (i in 1:length(genomes)) {
       #print(genomes[i])
       #i<-1
      genomes[[i]]<- mutate(genomes[[i]])
      #print(genomes[i])
    }
    gen <- runGeneration(genomes, x)
    gen.copy <- gen
    gen.copy$generation <- a
    gen.history <- rbind(gen.history, gen.copy)

  }

  top.scorer <- as.numeric(which.max(gen$fitness))
  final.weights<-genomes[[top.scorer]]
  return(list(best.agent.weights = final.weights, gen.history=gen.history))
}




```
Attempt to Plot
```{r}
gridBotANN.PlotState <-function(state){
    gridVisPlot<-state[[2]]
    df<- as.data.frame(which(gridVisPlot == 1, arr.ind = T))
    df$type <- 1
    df2<- as.data.frame(which(gridVisPlot == 0, arr.ind = T))
    df2$type <- 0
    df.all<- rbind(df, df2)
    df.all$type <- factor(df.all$type)
    p <- ggplot(df.all, aes(x=col, y=row, shape = type))+
      scale_shape_manual(values = c(21,22), guide=F)+
      geom_point(size=5)+
      scale_y_reverse()+
      theme_minimal()+
      theme(axis.text = element_blank(),
            panel.grid=element_blank(),
            axis.title= element_blank())
    return(p)
    }

```
